Fractional Types

### 2016-12-18 typespp.tex
pointed combinators

\newcommand{\Rule}[4]{
\makebox{{\rm #1}
$\displaystyle
\frac{\begin{array}{l}#2\\\end{array}}
{\begin{array}{l}#3\\\end{array}}$
 #4}}
\newcommand{\proves}{\vdash}
\newcommand{\jdg}[3]{{#1} #3}
\newcommand{\evalone}[2]{#1~\triangleright~#2}
\newcommand{\evaloneb}[2]{#1~\triangleleft~#2}
\newcommand{\unitv}{()}
\newcommand{\unitepl}{\mathsf{{unite_+l}}}
\newcommand{\unitipl}{\mathsf{{uniti_+l}}}
\newcommand{\unitepr}{\mathsf{{unite_+r}}}
\newcommand{\unitipr}{\mathsf{{uniti_+r}}}
\newcommand{\swapp}{\mathsf{{swap_+}}}
\newcommand{\assoclp}{\mathsf{{assocl_+}}}
\newcommand{\assocrp}{\mathsf{{assocr_+}}}
\newcommand{\unitetl}{\mathsf{{unite_{\star}l}}}
\newcommand{\unititl}{\mathsf{{uniti_{\star}l}}}
\newcommand{\unitetr}{\mathsf{{unite_{\star}r}}}
\newcommand{\unititr}{\mathsf{{uniti_{\star}r}}}
\newcommand{\swapt}{\mathsf{{swap_{\star}}}}
\newcommand{\assoclt}{\mathsf{{assocl_{\star}}}}
\newcommand{\assocrt}{\mathsf{{assocr_{\star}}}}
\newcommand{\absorbr}{\mathsf{{absorbr}}}
\newcommand{\absorbl}{\mathsf{{absorbl}}}
\newcommand{\factorzr}{\mathsf{{factorzr}}}
\newcommand{\factorzl}{\mathsf{{factorzl}}}
\newcommand{\dist}{\mathsf{{dist}}}
\newcommand{\factor}{\mathsf{{factor}}}
\newcommand{\distl}{\mathsf{{distl}}}
\newcommand{\factorl}{\mathsf{{factorl}}}
\newcommand{\idiso}{\mathsf{{id}}}

@@

\begin{figure}[t]
\[\begin{array}{rrcll}
\unitetl/ :&  \order{\idiso} \circledast \mathbb{T} & \leftrightspoon & \mathbb{T} &: \unititl/ \\
\unitetr/ :&  \mathbb{T} \circledast \order{\idiso} & \leftrightspoon & \mathbb{T} &: \unititr/ \\
\swapt/ :&  \mathbb{T}_1 \circledast \mathbb{T}_2 & \leftrightspoon & \mathbb{T}_2 \circledast \mathbb{T}_1 &: \swapt/ \\
\assoclt/ :&  \mathbb{T}_1 \circledast (\mathbb{T}_2 \circledast \mathbb{T}_3) & \leftrightspoon & (\mathbb{T}_1 \circledast \mathbb{T}_2) \circledast \mathbb{T}_3
  &: \assocrt/ \\
\\
\eta-_k :& [\order{\idiso},\idiso] & \leftrightblackspoon &
  [\iorder{c},1/c^k] \circledast [\order{c},c^k] &: \epsilon-_k \\
\eta+ :& \order{\idiso} & \leftrightspoon & \order{c} \circledast \iorder{c} &: \epsilon+ \\
\end{array}\]

Each combinator $\rho$ has an inverse.
\caption{$\Pi^/$ fractional combinators (dependent version)}
\label{pifracd:comb}
\end{figure}

@@

The first step is to enrich the types to pointed types $\mathbb{T_\bullet}$
where each type $\mathbb{T}$ refers to a particular value $v$ in
$\mathbb{T}$. We then allow indices of values $k$ to be range over a collection
of types. In particular, the type
$\exists k.[\order{c} \circledast \iorder{c}, \langle c^k, 1/c^k \rangle]$
imposes a constraint on the two values associated with the types $\order{c}$ and
$\iorder{c}$; they must be ``synchronized'' sharing the same $k$.

>>>>

\newcommand{\Rule}[4]{
\makebox{{\rm #1}
$\displaystyle
\frac{\begin{array}{l}#2\\\end{array}}
{\begin{array}{l}#3\\\end{array}}$
 #4}}
\newcommand{\proves}{\vdash}
\newcommand{\jdg}[3]{{#1} #3}
\newcommand{\evalone}[2]{#1~\triangleright~#2}
\newcommand{\evaloneb}[2]{#1~\triangleleft~#2}
\newcommand{\unitv}{()}
\newcommand{\synch}{\mathsf{{synch}}}
\newcommand{\unitepl}{\mathsf{{unite_+l}}}
\newcommand{\unitipl}{\mathsf{{uniti_+l}}}
\newcommand{\unitepr}{\mathsf{{unite_+r}}}
\newcommand{\unitipr}{\mathsf{{uniti_+r}}}
\newcommand{\swapp}{\mathsf{{swap_+}}}
\newcommand{\assoclp}{\mathsf{{assocl_+}}}
\newcommand{\assocrp}{\mathsf{{assocr_+}}}
\newcommand{\unitetl}{\mathsf{{unite_{\star}l}}}
\newcommand{\unititl}{\mathsf{{uniti_{\star}l}}}
\newcommand{\unitetr}{\mathsf{{unite_{\star}r}}}
\newcommand{\unititr}{\mathsf{{uniti_{\star}r}}}
\newcommand{\swapt}{\mathsf{{swap_{\star}}}}
\newcommand{\assoclt}{\mathsf{{assocl_{\star}}}}
\newcommand{\assocrt}{\mathsf{{assocr_{\star}}}}
\newcommand{\absorbr}{\mathsf{{absorbr}}}
\newcommand{\absorbl}{\mathsf{{absorbl}}}
\newcommand{\factorzr}{\mathsf{{factorzr}}}
\newcommand{\factorzl}{\mathsf{{factorzl}}}
\newcommand{\dist}{\mathsf{{dist}}}
\newcommand{\factor}{\mathsf{{factor}}}
\newcommand{\distl}{\mathsf{{distl}}}
\newcommand{\factorl}{\mathsf{{factorl}}}
\newcommand{\idiso}{\mathsf{{id}}}

@@

\begin{figure}[t]
{\footnotesize
\[\begin{array}{rrcll}
\unitetl/ :&  \exists\overline{k}.[\order{\idiso} \circledast \mathbb{T},
                                                   \langle \idiso,v \rangle]
  & \leftrightblackspoon &
                    \exists\overline{k}.[\mathbb{T},v]
  &: \unititl/ \\
\unitetr/ :&  \exists\overline{k}.[\mathbb{T} \circledast \order{\idiso},
                                                    \langle v,\idiso]
  & \leftrightblackspoon &
                    \exists\overline{k}.[\mathbb{T},v]
  &: \unititr/ \\
\swapt/ :&  \exists\overline{k}.[\mathbb{T}_1 \circledast \mathbb{T}_2,
                                                   \langle v_1,v_2 \rangle]
 & \leftrightblackspoon &
                   \exists\overline{k}.[\mathbb{T}_2 \circledast \mathbb{T}_1,
                                                  \langle v_2,v_1 \rangle]
  &: \swapt/ \\
  \assoclt/ :&  \exists\overline{k}.[\mathbb{T}_1 \circledast (\mathbb{T}_2 \circledast \mathbb{T}_3),
                                        \langle v_1,\langle v_2,v_3 \rangle]
  & \leftrightblackspoon &
                       \exists\overline{k}.[(\mathbb{T}_1 \circledast \mathbb{T}_2) \circledast \mathbb{T}_3,
                                       \langle \langle v_1,v_2 \rangle , v_3 \rangle]
  &: \assocrt/ \\
\\
\eta- :& [\order{\idiso},\idiso] & \leftrightblackspoon &
  \exists{k}.[\iorder{c} \circledast \order{c}, \langle 1/c^k, c^k \rangle] &: \epsilon- \\
\eta+ :& [\order{\idiso},\idiso] & \leftrightblackspoon &
  \exists{k}.[\order{c} \circledast \iorder{c}, \langle c^k,1/c^k \rangle] &: \epsilon+ \\
\end{array}\]
\[\begin{array}{c}
\Rule{}
{}
{\jdg{}{}{\idiso/ : \exists\overline{k}.[\mathbb{T},v] \leftrightblackspoon
  \exists\overline{k}.[\mathbb{T},v]}}
{}
~~
\Rule{}
{\jdg{}{}{\alpha : c_1 \isotwo c_2}}
{\jdg{}{}{(\order{\alpha})_k :
 [\order{c_1},c_1^k]  \leftrightblackspoon
 [\order{c_2}.c_2^k]}}
{}
\\
\\
\Rule{}
{\jdg{}{}{\rho_1 : \exists{\overline{k_1}}.
  [\mathbb{T}_1,v_1] \leftrightblackspoon [\mathbb{T}_2,v_2]} \quad
  \rho_2 : \exists{\overline{k_2}}.
  [\mathbb{T}_2,v_2] \leftrightblackspoon [\mathbb{T}_3,v_3]}
{\jdg{}{}{\rho_1 \circledcirc \rho_2 :
 \exists{\overline{k_1}\overline{k_2}}.
 [\mathbb{T}_1,v_1] \leftrightblackspoon [\mathbb{T}_3},v_3]}
{}
\\
\\
\Rule{}
{\jdg{}{}{\rho_1 :
  \exists{\overline{k_1}}.[\mathbb{T}_1,v_1] \leftrightblackspoon
  \exists{\overline{k_2}}.[\mathbb{T}_2,v_2]}
 \qquad \rho_2 :
  \exists{\overline{k_3}}.[\mathbb{T}_3,v_3] \leftrightblackspoon
  \exists{\overline{k_4}}.[\mathbb{T}_4,v_4]}
{\jdg{}{}{\rho_1 \circledast \rho_2 :
  \exists{\overline{k_1k_2k_3k_4}}.
  [\mathbb{T}_1 \circledast \mathbb{T}_3, \langle v_1,v_3 \rangle]
  \leftrightblackspoon
  [\mathbb{T}_2 \circledast \mathbb{T}_4}, \langle v_2,v_4 \rangle]}
{}
\\
\\
\Rule{}
{\jdg{}{}{}}
{\jdg{}{}{\synch : \exists{\overline{k}k'}.[\mathbb{T},v]
  \leftrightblackspoon
   [\mathbb{T},v[k'\mapsto ??]]}}
{}
\end{array}\]
Each combinator $\rho_\bullet$ has an inverse.
}
\caption{$\Pi^/$ fractional combinators (dependent version)}
\label{pifracd:comb}
\end{figure}

@@

The first step is to enrich the types to pointed types $\mathbb{T_\bullet}$
where each type $\mathbb{T}$ refers to a particular value $v$ in
$\mathbb{T}$. We then allow indices of values $k$ to be range over a collection
of types. In particular, the type
$\exists k.[\order{c} \circledast \iorder{c}, \langle c^k, 1/c^k \rangle]$
imposes a constraint on the two values associated with the types $\order{c}$ and
$\iorder{c}$; they must be ``synchronized'' sharing the same $k$. In general, a
pointed type will be of the form $\exists\overline{k}.[\mathbb{T},v]$ for some
sequence of integers $\overline{k}$ and some value $v$ in $\mathbb{T}$ which
might of course refer to the integers $\overline{k}$ to enforce some
synchronization among its sub-values.

Here is how the circuit in Fig.~\ref{fig:zigzag} evaluates in the forward and
backwards directions:
\[\begin{array}{rcl}
[\order{c},c^i]
&\stackrel{\unitetl/}{\longmapsto}&
  [\order{id}\circledast\order{c}, \langle \idiso,c^i \rangle] \\
&\stackrel{\eta+\circledast\idiso/}{\longmapsto}&
  \exists{k}.[(\order{c}\circledast\iorder{c})\circledast\order{c},
  \langle \langle c^k,1/c^k \rangle,c^i \rangle] \\
&\stackrel{\assocrt/}{\longmapsto}&
  \exists{k}.[\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^k, \langle 1/c^k,c^i \rangle\rangle] \\
&\stackrel{\synch}{\longmapsto}&
  [\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^i,c^i \rangle\rangle] \\
&\stackrel{\idiso/\circledast\epsilon-}{\longmapsto}&
  [\order{c}\circledast\order{\idiso},
  \langle c^i, \idiso \rangle] \\
&\stackrel{\unitetr/}{\longmapsto}&
  [\order{c}, c^i]
\end{array}\]

\[\begin{array}{rcl}
  [\order{c}, c^i]
&\stackrel{\unitetr/}{\longmapsfrom}&
  [\order{c}\circledast\order{\idiso},
  \langle c^i, \idiso \rangle] \\
&\stackrel{\idiso/\circledast\epsilon-}{\longmapsfrom}&
  \exists{k}.[\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^k,c^k \rangle\rangle] \\
&\stackrel{\synch}{\longmapsfrom}&
 [\order{c}\circledast(\iorder{c}\circledast\order{c}),
  \langle c^i, \langle 1/c^i,c^i \rangle\rangle] \\
&\stackrel{\assocrt/}{\longmapsfrom}&
  [(\order{c}\circledast\iorder{c})\circledast\order{c},
  \langle \langle c^i,1/c^i \rangle,c^i \rangle] \\
&\stackrel{\eta+\circledast\idiso/}{\longmapsfrom}&
  [\order{c}\circledast\order{\idiso}, \langle c^i, \idiso] \\
&\stackrel{\unitetl/}{\longmapsfrom}&
  [\order{c},c^i]
\end{array}\]
